Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-10 paper/disc20clustering.tex
Comments

The algorithm in this section, and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from \cite{repartition-disc}.
The difference is in the the choice of partition after a component merge.
In DET, the partition was arbitrary.
In the algorithm for the learning model, we choose the partition closest to the initial partition.
In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).

\maciek{The remainder of the paper used "configuration C" to denote the partition P. Easiest to adjust here. \mahmoud{are "configuration" and "partition" the same thing everywhere? Then let's use "partition" only.}}

@@

With each repartitioning,
\PPL moves to a partition that minimizes the distance to the initial partition $P_I$.
As a result,
\PPL never moves to a~partition that is more than $\Delta^*$  node migrations away from $P_I$.
This invariant latter ensures us that \PPL does not pay too much while recovering $P_F$.
We emphasis that a repartitioning by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
Meaning, it never moves to a partition beyond distance $\Delta^*$.

@@

\begin{theorem}	\label{thm:upperbound}
	\PPL reaches the final partition $P_F$
	 and it is $(2\cdot (k-1)\cdot\ell)$-competitive.
\end{theorem}
\begin{proof}
	On each inter-cluster request,
	the algorithm enumerates all $\ell$-way partitions of components
	that are in the same (closest) distance to $P_I$.
	That is, 
	once it reaches a partition $P$ at distance $\Delta^* = \Delta(P)$,
	it does not move to a partition
	$P', \Delta(P') > \Delta^*$,
	before it enumerates all partitions at distance $\Delta^*$.
	Therefore,
	\PPL eventually reaches  the partition $P_F$ at distance $\Delta^*=\OPT$.
	With each distinct request, 
	the size of some component increases by one.
	For any cluster $F_i \in P_F$,
	we have $\sum_{C \in F_i} |C| = k$.
	A component $C \in F_i$,
	 initially begins as an isolated node and it grows by gaining $|C|-1$ more nodes.
	Hence, the total number of times a component in $F_i$ grows is 
	$\sum_{C \in F_i} (|C|-1) \leq k-1$.
	Therefore, there are at most  $(k-1)\cdot\ell $ distinct requests
	for which \PPL performs a repartitioning.
	I.e.,
	 \PPL performs at most $(k-1)\cdot\ell $ repartitionings.
	By Lemma \ref{lemma:rebalancecost},
	each repartitioning costs at most $2\cdot\OPT$.
	The total cost is therefore at most $2\cdot\OPT\cdot (k-1) \cdot\ell$, which implies the competitive ratio.
\end{proof}

>>>>

The algorithm in this section, and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from \cite{repartition-disc}.
The difference is in the the choice of partition after a component merge.
In DET, the partition was arbitrary.
In the algorithm for the learning model, we choose the partition closest to the initial partition.
\maciek{component respecting?}
In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).

\maciek{The remainder of the paper used "configuration C" to denote the partition P. Easiest to adjust here. \mahmoud{are "configuration" and "partition" the same thing everywhere? Then let's use "partition" only.\maciek{I thought partition is on component level, and configuration on node level (todo check where to add component-respecting to the partition in this chapter). Maybe say in the intro in this chapter that configuration = partition? We use both heavily.}}}

@@

With each repartitioning,
\PPL moves to a \maciek{component-respecting} partition that minimizes the distance to the initial partition $P_I$.
As a result,
\PPL never moves to a~partition that is more than $\Delta^*$  node migrations away from $P_I$.
This invariant latter ensures us that \PPL does not pay too much while recovering $P_F$.
We emphasis that a repartitioning by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
Meaning, it never moves to a partition beyond distance $\Delta^*$.

@@

\begin{theorem}	\label{thm:upperbound}
	\PPL reaches the final partition $P_F$
	 and it is $(2\cdot (k-1)\cdot\ell)$-competitive.
\end{theorem}
\begin{proof}
	On each inter-cluster request,
	the algorithm enumerates all \maciek{component-respecting} $\ell$-way partitions of components
	that are in the same (closest) distance to $P_I$.
	That is, 
	once it reaches a partition $P$ at distance $\Delta^* = \Delta(P)$,
	it does not move to a partition
	$P', \Delta(P') > \Delta^*$,
	before it enumerates all partitions at distance $\Delta^*$.
	Therefore,
	\PPL eventually reaches  the partition $P_F$ at distance $\Delta^*=\OPT$.
	With each distinct request, 
	the size of some component increases by one.
	For any cluster $F_i \in P_F$,
	we have $\sum_{C \in F_i} |C| = k$.
	A component $C \in F_i$,
	 initially begins as an isolated node and it grows by gaining $|C|-1$ more nodes.
	Hence, the total number of times a component in $F_i$ grows is 
	$\sum_{C \in F_i} (|C|-1) \leq k-1$.
	Therefore, there are at most  $(k-1)\cdot\ell $ distinct requests
	for which \PPL performs a repartitioning.
	I.e.,
	 \PPL performs at most $(k-1)\cdot\ell $ repartitions.
	By Lemma \ref{lemma:rebalancecost},
	each repartitioning costs at most $2\cdot\OPT$.
	The total cost is therefore at most $2\cdot\OPT\cdot (k-1) \cdot\ell$, which implies the competitive ratio.
\end{proof}
